---
title: "Zip Competition"
author: "Jingni Pang & Zhaoxin Zhang"
output:
  html_document:
    df_print: paged
always_allow_html: true
editor_options:
  chunk_output_type: console
  markdown:
    wrap: sentence
---

GitHub Repository: https://github.com/ZZhang727/Pang_Zhang_ZIP

# Exective Summary

This report is written to record the process of using R to leverage this historical data (from January 2005 to June 2011) to forecast daily electricity demand for July 2011. We applied different models to forecast electricity demand in July 2010 and compared the forecasting results with the actual data to identify the most accurate model. The model using Prophet + TBATS had the best results. Therefore, we applied the Prophet + TBATS model to forecast electricity load in July 2011.


```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

# Section 1: Data Wrangling

```{r include=FALSE}
# Install Packages
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(zoo)
library(kableExtra) 
```

## 1.1 Import and Convert Data
Hourly electricity load data spanning January 2005 to June 2011 were used to forecast daily load for July 2011. As the raw data consist of 24 hourly observations per day, daily mean load values were computed by averaging across all hours within each calendar day, yielding a single daily aggregate series. The resulting series was then converted to a time series object for subsequent modeling.
```{r include=FALSE}

install.packages("readxl")
library(readxl)

#import load data 

load <- read_excel(path  = "../ZIP_Competition/load.xlsx", sheet = 1, col_names = TRUE)
class(load$date)

#Convert to daily data

load_daily <- load %>%
  mutate(date = as.Date(date)) %>%                  
  mutate(daily_mean = rowMeans(select(., h1:h24), 
                               na.rm = TRUE)) %>%
  select(meter_id, date, daily_mean) 

#Convert to time series
load_ts <- msts(load_daily$daily_mean,
                seasonal.periods = c(7, 365.25),
                start = c(2005, 1))
```

```{r echo=FALSE}
autoplot(load_ts) +
  labs(
    title = "Graph 1. Daily Mean Electricity Load (2005–2011)",
    x = "Time",
    y = "Mean Load (kW)"
  ) +
  theme_minimal()
```

# 2. Methodology
The data were divided into a training set and a held-out test set. The training set spans January 2005 through June 2010 and was used to estimate model parameters and capture underlying temporal structure, including trend and seasonal components. The test set comprises the final 31 days of the sample (July 2010), which was withheld from model fitting and used exclusively for forecast evaluation, enabling comparison of predicted against realized daily mean load values across models. 
```{r include=FALSE}
#Use data from January 1, 2005, to July 31, 2010

load_daily_2010July <- load_daily %>%
  filter(date <= as.Date("2010-07-31"))

load_daily_2010July <- msts(load_daily_2010July$daily_mean,
                     seasonal.periods = c(7, 365.25),
                     start = c(2005, 1))
n_for <- 31

#Train the data
load_ts_train <- subset(load_daily_2010July, end = length(load_daily_2010July) - n_for)
print(load_ts_train,10)
#Test
load_ts_test  <- subset(load_daily_2010July, start = length(load_daily_2010July) - n_for + 1)


```

```{r echo=FALSE}

autoplot(load_ts_train)+
    labs(
    title = "Graph 2. Daily Mean Electricity Load (Jan 2005– June 2010)",
    x = "Time",
    y = "Mean Load (kW)"
  ) +
  theme_minimal()

autoplot(load_ts_test)+
    labs(
    title = "Graph 3. Daily Mean Electricity Load (July 2011)",
    x = "Time",
    y = "Mean Load (kW)"
  ) +
  theme_minimal()
```


# 3. Testing
To evaluate forecast accuracy, each candidate model was trained on daily mean load data from January 2005 through June 2010 and used to generate out-of-sample forecasts for July 2010, which were then compared against realized values using RMSE, MAE, and MAPE. A Seasonal Naïve model was estimated as a baseline; however, it produced forecast errors, indicating that the data exhibit temporal patterns beyond simple seasonal repetition. Four additional specifications were subsequently evaluated: TBATS, Neural Network (NNETAR with Fourier terms), a TBATS–Neural Network combination, Prophet, and a TBATS–Prophet combination. Across all accuracy metrics, the TBATS–Prophet combination yielded the lowest errors and was selected as the final forecasting model to predict the electricity load in July 2011.

```{r echo=FALSE}
# Model1 Naive
SNAIVE_test <- snaive(load_ts_train, h = n_for, holdout = FALSE)
SNAIVE_scores<- accuracy(SNAIVE_test$mean, load_ts_test)
SNAIVE_scores
```


```{r include=FALSE}
# Test TBATS
TBATS_fit <- tbats(load_ts_train)
TBATS_for <- forecast(TBATS_fit, h = n_for)
TBATS_scores <- accuracy(TBATS_for$mean, load_ts_test)
```

```{r include=FALSE}
#Neural Network
NN_fit <- nnetar(load_ts_train,
                 p = 7,
                 P = 1,
                 xreg = fourier(load_ts_train, K = c(2, 2)))

NN_for <- forecast(NN_fit,
                   h = n_for,
                   xreg = fourier(load_ts_train, K = c(2, 2), h = n_for))

NN_scores <- accuracy(NN_for$mean, load_ts_test)
```

```{r include=FALSE}
# Combination: TBATS + NN
COMBO_for <- (TBATS_for$mean + NN_for$mean) / 2
COMBO_scores <- accuracy(COMBO_for, load_ts_test)
print(COMBO_scores)
```

```{r include=FALSE}
#Prophet
install.packages("prophet")
library(prophet)

prophet_df <- load_daily %>%
  select(date, daily_mean) %>%
  rename(ds = date, y = daily_mean) %>%
  filter(ds <= as.Date("2010-06-30"))

prophet_fit <- prophet(prophet_df,
                       weekly.seasonality = TRUE,
                       yearly.seasonality = TRUE,
                       daily.seasonality = FALSE)

future <- make_future_dataframe(prophet_fit, periods = 31)
prophet_for <- predict(prophet_fit, future)

prophet_pred <- tail(prophet_for$yhat, 31)

prophet_scores <- accuracy(prophet_pred, load_ts_test)
print(prophet_scores)
```

```{r include=FALSE}
# Combination: TBATS + Prophet
COMBO_TP_for <- (TBATS_for$mean + prophet_pred) / 2
COMBO_TP_scores <- accuracy(COMBO_TP_for, load_ts_test)
```

## Compare Results
```{r echo=FALSE}
scores <- as.data.frame(rbind(TBATS_scores,NN_scores,COMBO_scores, prophet_scores,COMBO_TP_scores))
row.names(scores) <- c("TBAT", "NN", "TBATS+NN", "Prophet", "TBAT+Prophet")
print(scores)
```
Among the five models evaluated, TBATS achieved the lowest RMSE (578.17), MAE (444.73), and MAPE (11.34%), indicating strong overall predictive accuracy. However, its ME of -37.89 suggests a systematic tendency to underestimate load. The TBATS–Prophet combination, while marginally higher on RMSE (584.26) and MAE (457.26), produced a ME of -9.07, which was the closest one to zero across all models. Therefore, the TBATS–Prophet combination was selected as the final model.

# Forecast Electricity Load in July 2011
To generate the final forecast for July 2011, both models (TBATS and Prophet) were re-estimated on the full available training data. The Prophet model was fitted on daily mean load data spanning January 2005 through June 2011, with weekly and yearly seasonality enabled and daily seasonality disabled. A 31-day future dataframe was constructed using make_future_dataframe(), and point forecasts were extracted from the predicted yhat values corresponding to July 2011. In parallel, a TBATS model was re-fitted on the complete daily mean load time series (load_ts) and used to generate a 31-day ahead forecast. The final combined forecast was computed as a weighted average of the two models, assigning a weight of 0.8 to Prophet and 0.2 to TBATS (the weight was selected after trying different weight combinations to find the most accurate result), yielding daily predicted load values for each day in July 2011.
```{r include=FALSE}
prophet_df_full <- load_daily %>%
  select(date, daily_mean) %>%
  rename(ds = date, y = daily_mean) %>%
  filter(ds <= as.Date("2011-06-30"))

prophet_fit_final <- prophet(prophet_df_full,
                             weekly.seasonality = TRUE,
                             yearly.seasonality = TRUE,
                             daily.seasonality = FALSE)

future_final <- make_future_dataframe(prophet_fit_final, periods = 31)
prophet_for_final <- predict(prophet_fit_final, future_final)

prophet_pred_final <- tail(prophet_for_final$yhat, 31)
TBATS_fit_final <- tbats(load_ts)
TBATS_for_final <- forecast(TBATS_fit_final, h = 31)

combo_prophet_tbats2 <- (prophet_pred_final * 0.8 + as.numeric(TBATS_for_final$mean) * 0.2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Export
submission_combo_pt2 <- data.frame(
  date = seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day"),
  load = as.numeric(combo_prophet_tbats2)
)

knitr::kable(submission_combo_pt2, 
             caption = "Predicted Daily Load for July 2011 (TBATS + Prophet)",
             col.names = c("Date", "Predicted Load"),
             digits = 3)
```

```{r include=FALSE}
#Export as csv.
write.csv(submission_combo_pt2, "./submission_combo_pt2.csv", row.names = FALSE)
```

#(When implementing Prophet and the TBATS–Prophet combination, we utilized Claude as a coding assistant to help develop and refine our code.)
